Malletwand: the Pendulum as a Handheld Interface to Musical Timing

### 2024-02-08 paper/paper.tex
Minor fixes

\subsection{Estimation Algorithm}
All processing are done on the microcontroller device. The microcontroller we used, STM32G031, is an Arm Cortex-M0+ processor without a dedicated FPU, but our complete algorithm runs well within its 64 MHz system clock.

@@

\subsubsection{Magnetometer}
\label{appendix:calib-mag}
This task can be formulated as follows: given a set of 3-dimensional vectors $\{ \tilde{\mathbf{m}}_{1 \dots n} \}$, find an affine transform $(\mathbf{Q}, \mathbf{o})$ such that the set of $\{\mathbf{Q} \tilde{\mathbf{m}}_i + \mathbf{o}\}$ best fits onto a unit sphere.

@@

Since the ALS estimator only requires an accumulated matrix $\mathbf{\Psi}$ and not the individual $\tilde{\mathbf{m}}_i$, we are able to carry out the calculation online in constant rather than linear space.

\subsection{Estimation}
\subsubsection{Facing}
\label{appendix:est-facing}
Our complementary filter design follows that in~\cite{Min_Complementary}, namely, a filter satisfying the transfer equation
\begin{equation}
\begin{split}
& \hat{\theta} = \frac 1 {s^2 + k_\mathrm{p} s + k_\mathrm{i}} \left( s \dot\theta + (k_\mathrm{p} s + k_\mathrm{i}) \theta \right) \text{,} \\
& k_\mathrm{p} = \frac {\omega_0} {Q_\mathrm{P}},\qquad k_\mathrm{i} = \omega_0^2\text{.}
\end{split}
\end{equation}
This filter balances between a proportional signal $\theta$ and a derivative signal $\dot\theta$. Here we treat the 2-dimensional orientation on the $xy$ plane as the value being estimated. The calibrated magnetometer provides the $\theta$ signal, while the calibrated gyroscope provides $\dot\theta$.

@@

\subsubsection{Motion}
\label{appendix:est-motion}
In the extended Kalman filter we formulate the state vector as $\mathbf{x} = \begin{bmatrix} \omega & A & \theta & B & \phi \end{bmatrix}^\mathsf{T}$, where the observation $\mathbf{z} = \begin{bmatrix} A \cos \theta & B \cos (\theta + \phi) \end{bmatrix}^\mathsf{T}$ is the gyroscope readings adjusted by the facing estimation,
\begin{equation}
\begin{split}
\hat{\mathbf{g}}' &= \mathbf{q}_{z, \theta} \hat{\mathbf{g}} \mathbf{q}_{z, \theta}^{-1} \\
\mathbf{q}_{z, \theta} &= \begin{bmatrix} 0 & 0 & \sin \frac \theta 2 & \cos \frac \theta 2 \end{bmatrix}
\text{,}
\end{split}
\end{equation}
projected onto the $xy$ plane. Varying $\theta$ over $[0, 2\pi)$ traces out an elliptic trajectory of this point and $\theta$ is the current phase within the period. $\omega$ is the rate of increment of $\theta$.

>>>>

\subsection{Estimation Algorithm}
All processing is done on the microcontroller device. The microcontroller we used, STM32G031, is an Arm Cortex-M0+ processor without a dedicated FPU, but our complete algorithm runs well within its 64 MHz system clock with software floating-point arithmetic.

@@

\subsubsection{Magnetometer}
\label{appendix:calib-mag}
This task can be formulated as follows: given a set of 3-dimensional vectors $\{ \tilde{\mathbf{m}}_i \}$, find an affine transform $(\mathbf{Q}, \mathbf{o})$ such that the set of $\{\mathbf{Q} \tilde{\mathbf{m}}_i + \mathbf{o}\}$ best fits onto a unit sphere.

@@

Since the ALS estimator only requires an accumulated matrix $\mathbf{\Psi}$ and not the individual $\tilde{\mathbf{m}}_i$, we are able to carry out the calculation online in constant rather than linear memory space.

\subsection{Estimation}
\subsubsection{Facing}
\label{appendix:est-facing}
Our complementary filter design follows that in~\cite{Min_Complementary}, namely, one satisfying the transfer equation
\begin{equation}
\begin{split}
& \hat{\theta} = \frac 1 {s^2 + k_\mathrm{p} s + k_\mathrm{i}} \left( s \dot\theta + (k_\mathrm{p} s + k_\mathrm{i}) \theta \right) \text{,} \\
& k_\mathrm{p} = \frac {\omega_0} {Q_\mathrm{P}},\qquad k_\mathrm{i} = \omega_0^2\text{.}
\end{split}
\end{equation}
This filter balances between a proportional signal $\theta$ and a derivative signal $\dot\theta$. Here we treat the 2-dimensional orientation on the $xy$ plane as the value being estimated. The calibrated magnetometer provides the $\theta$ signal, while the calibrated gyroscope provides $\dot\theta$.

@@

\subsubsection{Motion}
\label{appendix:est-motion}
In the extended Kalman filter we formulate the state vector as $\mathbf{x} = \begin{bmatrix} \omega & A & \theta & B & \phi \end{bmatrix}^\mathsf{T}$, where the observation $\mathbf{z} = \begin{bmatrix} A \cos \theta & B \cos (\theta + \phi) \end{bmatrix}^\mathsf{T}$ is the gyroscope readings adjusted by the facing estimation,
\begin{equation}
\begin{split}
\hat{\mathbf{g}}' &= \mathbf{r}_{z, \theta} \hat{\mathbf{g}} \mathbf{r}_{z, \theta}^{-1} \\
\mathbf{r}_{z, \theta} &= \begin{bmatrix} 0 & 0 & \sin \frac \theta 2 & \cos \frac \theta 2 \end{bmatrix}
\text{,}
\end{split}
\end{equation}
projected onto the $xy$ plane. Varying $\theta$ over $[0, 2\pi)$ traces out an elliptic trajectory of this point and $\theta$ is the current phase within the period. $\omega$ is the rate of increment of $\theta$.
